---
title: "Atmosphere and Weather Lab 3: Radiation"
runtime: shiny
output: html_document
---
### Planck's Law and Wien's Law

```{r, echo = F, message=F}
library(tidyverse)
library(ggtext)
shinyjs::useShinyjs()
```

**Planck's law** says that the energy spectrum ($E(\lambda)$) of a black body is related to its absolute temperature ($T$), and is given by

$$E(\lambda) = \frac{C_{1}\lambda^{-5}}{\textrm{EXP}(\frac{C_2}{\lambda T})-1},$$
where $C_1$ = 3.74x10^8^ Wm^-2^, $C_2$ = 1.44x10^4^ $\mu mK$, $\lambda$ is the wavelength in $\mu m$, and $T$ is temperature in K. 

**Wien's law** says that the wavelength of maximum emission for a black body ($\lambda_m$) is inversely proportional to its temperature. Wien's law is given by

$$\lambda_{m} = \frac{2897\mu mK}{T},$$
where $T$ is temperature in K.

### Visualizing Planck's Law and Wien's Law

```{r, echo = F}
e_lam <- function(c1, c2, lambda, t){
  c1 * lambda ^ (-5) * exp(-1 * (c2/ (lambda * t) ))
}
w_law <- function(t){
  2897/t
}

c1 <- 3.74e8
c2 <- 1.44e4
earth_irrad <- tibble(irrad = 26,
                      lambda = 10^1,
                      label = "Irradiance of\n the Earth")
sun_irrad <- tibble(irrad = 83468929,
                      lambda = 0.7,
                      label = "Irradiance of\n the Sun",
                      hjust = 0,
                      vjust = 1)
```

Use the slider below to visualize how changing the absolute temperature of a black body changes its emission spectra. The wavelength of maximum emission, $\lambda_m$, is reported at the top of the figure. 

```{r echo = FALSE}
# shiny::textInput("lambda",shiny::HTML("Enter &lambda;"),
#                  value = "")
# shiny::textInput("T_",shiny::HTML("Enter T"),
#                  value = "5800")
shinyWidgets::chooseSliderSkin("Flat", color = "#112446")

shiny::sliderInput("T_",shiny::HTML("Select temperature"),
                   min = 100,
                   max = 5800,
                   step = 5,
                   value = 100)

shinyWidgets::awesomeCheckbox(
   inputId = "scale",
   label = shiny::HTML("Scale earth irradiance by 1e6 Wm<sup>-2</sup> &mu;m<sup>-1</sup>"), 
   value = F
)
```
 
```{r echo = FALSE}

df <- shiny::reactive({
        int <- tibble::tibble(lambda = seq(10^-1, 10^2, 10^-1),
                 irrad = e_lam(c1, c2, 
                               lambda = lambda,
                               t = as.numeric(input$T_)),
                 earth_irrad = e_lam(c1, c2, 
                               lambda = lambda,
                               t = 288)) 
        
        if (input$scale){
          int$earth_irrad <- int$earth_irrad * 1e6
          earth_irrad$irrad <- earth_irrad$irrad * 1e6
          earth_irrad$label <- "Scaled earth irradiance"
        }
        
        lmax <- tibble(lambda = w_law(input$T_),
                       irrad = max(int$irrad))
        
        return(list(int = int,
                    earth_irrad = earth_irrad,
                    lmax = lmax))
      })

renderPlot({
  if (input$T_ < 5800){
    
    p <- ggplot(df()$int) +
      geom_line(aes(x = lambda, y = irrad)) +
      geom_line(aes(x = lambda, y = earth_irrad),
                color = "purple") +
      geom_richtext(data = df()$earth_irrad,
                    aes(x = lambda, y = irrad, label = label)) +
      scale_x_continuous(trans = "log10") +
      theme_bw() +
      labs(y = "Irradiance (Wm<sup>-2</sup> &mu;m<sup>-1</sup>)",
           x = "Wavelength &lambda; (&mu;m)",
           title = paste0("Black body energy spectra (&lambda;<sub>m</sub>: ",
                         round(df()$lmax$lambda,3),")")) +
      theme(axis.title.y = element_textbox(orientation = "left-rotated"),
            plot.title = element_textbox(),
            axis.text.x = element_text(vjust = -2),
            axis.title.x = element_textbox()) +
      annotation_logticks(sides = "b",
                          outside = T) +
      coord_cartesian(clip = "off") +
      scale_y_continuous(expand = c(0.025,0.025))
   
  } else if (input$T_ == 5800)  {
    
    p <- ggplot(df()$int) +
      geom_line(aes(x = lambda, y = irrad)) +
      geom_line(aes(x = lambda, y = earth_irrad),
                color = "purple") +
      geom_richtext(data = sun_irrad,
                    aes(x = lambda, y = irrad, label = label,
                        hjust = hjust,
                        vjust = vjust)) +
      geom_richtext(data = df()$earth_irrad,
              aes(x = lambda, y = irrad, label = label)) +
      scale_x_continuous(trans = "log10") +
      theme_bw() +
      labs(y = "Irradiance (Wm<sup>-2</sup> &mu;m<sup>-1</sup>)",
           x = "Wavelength &lambda; (&mu;m)",
           title = paste0("Black body energy spectra (&lambda;<sub>m</sub>: ",
                         round(df()$lmax$lambda,3),")")) +
      theme(axis.title.y = element_textbox(orientation = "left-rotated"),
            plot.title = element_textbox(),
            axis.text.x = element_text(vjust = -2),
            axis.title.x = element_textbox()) +
      annotation_logticks(sides = "b",
                          outside = T) +
      coord_cartesian(clip = "off") +
      scale_y_continuous(expand = c(0.025,0.025))
    
  }
  p
})
```